StarCraft Bots and Competitions

### 2015-12-16 ECGG_StarCraftCompetitions/ECGG_StarCraftCompetitions.tex
Fix typos

Some bots  using divide-and-conquer, assume  that each of  the modules
can act independently  and that their actions can  be executed without
interference.  BBQ,  UAlbertaBot and  AIUR, however use  an arbitrator
({\em Game Commander'} in UAlbertaBot) that makes sure that modules do
not send contradictory orders to  the same unit.  However, very little
bots handle  the problem of  how to coordinate resource  usage amongst
modules, for  instance BTHAI uses a  first-come-first-serve policy for
spending resources,  the first module  that requests resources  is the
one that gets them. Nova and Skynet are exceptions, and implement some
rudimentary prioritization based on the high level strategy. Following
available  resources and  timing,  AIUR's {\em  Spend Manager}  orders
Base,  Production   and  Construction  Managers  what   they  have  to
build/produce. It also orders to start tech research and upgrades. The
idea here is not to let the different managers allocate the resources they want, but to do the opposite, that is: finding how the AI can
spend the available money.

@@

A total of 8 bots competed in the 2013 AIIDE competition with many of the same names from the 2012 competition, and some key updates were made to existing bots for 2013 which shook up the results from the previous years. A total of 5597 games were played during the 5 day tournament. UAlbertaBot took 1st place with a dominant 84.49\% win rate, with Skynet in 2nd with 72.77\%. The major addition to UAlbertaBot was a combat simulation package called SparCraft. UAlbertaBot was re-engineered so that it now always attacked from the first combat unit created, using the SparCraft system to simulate whether or not the bot could a fight against known enemy units, retreating automatically if it thought it could not. This addition, combined with some additional bug fixes led to the victory. Skynet and Aiur both implemented strategy learning in 2013, which is evident from \ref{fig:aiide2013} in which both Skynet and Aiur's win percentage over time climb dramatically from the first few rounds of the tournament to the later rounds. Ximp unfortunately crashed all games on the Fortress map and lost 10\% of its games for free as a result, which meant it could have possibly came in 2nd place if not for all those crashes.

@@

The results of the 2014 competition were dramatically different from previous years. Since 2011 the top 3 bots in the AIIDE competition had been some permutation of Skynet, UAlbertaBot, and Aiur, and in 2014 the top 3 finishers were all completely different. In first place was IceBot with a win rate of 85.85\%, which had a complete overhaul in terms of strategy and tactics from the previous year. IceBot had several strategies which it could implement, most starting with an early bunker defense and transitioning into attacking once a set of heuristics conditions had been met. IceBot's source code was originally based on Aiur's modular design, and changed to play the Terran race. In a close 2nd place finish was Ximp - which had fixed the crashes and logic errors which had plagued it the year before. Ximp implemented a solid forge-first fast expansion strategy which transitioned into late-game carrier play. By building a large amount of early photon cannons in its expansion, Ximp was easily able to stop most of the early rushing bots, and then holding them off until its carriers cleaned up in the late game. 3rd place was taken by LetaBot, a new Terran bot whose source code was written on top of the 2012 version of UAlbertaBot and changed to play the Terran race. LetaBot implemented a number of strategies including a 'cheesy' but strong bunker rush strategy which killed many bots very quickly, as well as a Wraith (mid-game flying unit) strategy.

@@

The 2015 AIIDE competition was the largest competition to date with 23 total bots competing. The competition environment was also changed so that the tournament was now run on virtual machines instead of physical machines. Doing this enabled the tournament to run for a full two weeks instead of the normal one week, resulting in a total of 20788 games being played - 90 full round robins between each bot pairing. 2015 saw the most even distribution of race selection ever in a Starcraft AI competition, with 5 new Zerg submissions along with 7 Protoss bots and 9 Terran bots. The first ever Random race entry was also submitted as UAlbertaBot had been updated to play all three of the available races. When playing Random race, Starcraft randomly assigns one of Protoss, Terran, or Zerg after the game has started, with your opponent not knowing which race you are until they have scouted you on the map. This provides a significant advantage as the enemy's initial build order must now have a period of uncertainty during which is has to guess what race it is playing against.

The results of the 2015 competition were extremely close, with 1st / 2nd place as well as 3rd / 4th place being separated by less than 1\% win rate. 1st place was won by Tscmoo, a new Zerg bot which played nearly a dozen different strategies, learning which one to choose over time via the file I/O system. In a close second place was ZZZKbot, which implement a 4-pool Zergling rush strategy every game. Despite the relatively simple strategy, most bots did not have proper defense capabilities and lost in very short games. In 3rd place was Overkill, another Zerg bot which had several different strategies including Mutalisks and Hydralisks, which learned over time as well. In a close 4th place was UAlbertaBot, which played Random and implemented three main rushing strategies - one for each race. An interesting result from this tournament was that despite UAlbertaBot coming 4th place, it finished with a winning rate greater than 50\% vs each other bot. Aiur came in 5th place and was a clear demonstration of how learning over time can dramatically improve results in a tournament, going from 63\% win rate early in the competition to a final win rate of over 73\%.

\begin{table}[!t]
\caption{Results of the five best bots of the AIIDE 2015 competition}
\label{tab:aiide2015}
\centering
\begin{tabular}{|c|c|c|}
\hline
{\bfseries Rank} & {\bfseries Bot} & {\bfseries Win \%} \\
\hline
1 & Tscmoo & 88.52\% \\
2 & ZZZKBot & 87.84\% \\
3 & Overkill & 80.69\% \\
4 & UAlbertaBot & 80.20\% \\
5 & Aiur & 73.02\% \\
\hline
\end{tabular}
\end{table}

@@

The two top bots of every bracket qualified
for the final round. Table~\ref{tab:cig-first-round} summarizes
the results.
Note that as BroodwarBotQ and BTHAI have the same number of wins,
their direct encounter was evaluated which accounted 6:4 for the BroodwarBotQ.
The bots going into the final were thus UAlbertaBot, Skynet (from bracket A)
and Xelnaga and BroodwarBotQ (from bracket B). All qualified bots play the
Protoss faction. Most bots proved pretty stable, only Xelnaga and Protoss 
Beast Jelly crashed relatively often (each in more than a quarter of the games). 
Crashing of course resulted in an instant win for the other bot.
In some cases, neither bot was able to finish the other off completely,
so that they went into a passive state. We manually ended such games after
around 15 minutes and assigned victory to the bot that had obtained more
points as indicated on the end game screen.

@@

The final round was played in a similar mode as each of the
first round brackets,
using another set of 5 previously unknown maps:
\emph{iCCup lost temple 2.4}, \emph{iCCup rush hour 3.1},
\emph{iCCup swordinthemoon 2.1}, \emph{iCCup yellow 1.1},
and \emph{La\_Mancha 1.1}. 
Letting each pairing play on each map twice again with
switching starting positions resulted in 30 games per bot.
The final results are displayed in table~\ref{tab:cig-final-round},
indicating Skynet as winner and UAlbertaBot as runner-up, being
almost equally strong, and the two other bots as clearly inferior.
The competition setup, documentation and results can be found
in\footnote{\url{http://ls11-www.cs.tu-dortmund.de/rts-competition/StarCraft-cig2011}}.

@@

\subsubsection{CIG 2014.}
The CIG 2014 competition was organized by Kyung-Joong Kim, Ho-Chul Cho, and In-Seok Oh of Sejong University. For 2014, the CIG competition used a total of 20 different maps which were unknown to the competitors before the competition started, which was by far the most maps ever used in a Starcraft AI competition and presented a challenge to many of the bots who entered. A total of 4680 games were played which meant that each bot played each other bot 60 times, or 3 times per map. The CIG 2014 competition was held just a few weeks before the AIIDE 2014 competition and so many of the entrants to the competition were identical, and the results definitely showed this. The top 4 bots were all the same as the 2014 AIIDE competition with IceBot coming 1st, Ximp in 2nd, LetaBot in 3rd and Aiur in 4th place. For descriptions of these bots please refer to the AIIDE 2014 competition description as none of the top-finishing bots had any significant changes made.

\begin{table}[t]
\caption{Results of the CIG 2014 competition.}
\label{tab:cig2014}
\centering
\begin{tabular}{|c|c|c|}
\hline
{\bfseries Rank} & {\bfseries Bot} & {\bfseries Win \%} \\
\hline
1 & IceBot & 83.1\% \\
2 & Ximp & 78.1\% \\
3 & LetaBot & 68.5\% \\
4 & Aiur & 66.1\% \\
5 & UAlbertaBot & 60.0\% \\ 
\hline
\end{tabular}
\end{table}

\subsubsection{CIG 2015.}
There were some significant rule changes to the CIG 2015 competition, which was once organized by mebers from Sejong University. The most significant rule change was that entrants no longer had to be open source, in an attempt to attract more competitors who may not want to open source their bots. The second rule change was that a single competitor could submit multiple entries to the competition - which caused some controversy among registrants since this introduces the possibility of collusion between entries. Thankfully no such collusion was detected during the competition. The 2015 competition was also not run quite as long as previous competitions with only 2730 games being played in total, or 30 between each bot pairing, 6 between each bot on each of the 5 chosen maps. There were several new Zerg entries to the CIG 2015 competition which ended up finishing in the top three positions, with results very similar to those of the AIIDE 2015 competition. ZZZBot took 1st place, tscmoo-Z (a Zerg bot written by tscmoo) came 2nd, and Overkill came 3rd. For descriptions of these bots and their strategies used please refer to the AIIDE 2015 competition section as the strategies remained largely unchanged between the CIG and AIIDE competitions.

@@

% As the figures show, the top three ranked bots in the competition (Skynet, Aiur and UalbertaBot) do not change their strategy at all depending on their opponent. For example, the Skynet bot (both in 2011 and 2012), always uses the same opening (\emph{two\_gates}), except when playing a Terran opponent, when it uses \emph{nony}. This reflects the trend that the performance of bots is still more dependent on carefully handcrafted and non-adaptive behaviors, than on on-line decision making procedures. This is so, since most of the problems that need to be solved in order to implement such procedures are still open.

>>>>

Some bots  using divide-and-conquer, assume  that each of  the modules
can act independently  and that their actions can  be executed without
interference.  BBQ,  UAlbertaBot and  AIUR, however use  an arbitrator
({\em Game Commander} in UAlbertaBot) that makes sure that modules do
not send contradictory orders to  the same unit.  However, very little
bots handle  the problem of  how to coordinate resource  usage amongst
modules, for  instance BTHAI uses a  first-come-first-serve policy for
spending resources,  the first module  that requests resources  is the
one that gets them. Nova and Skynet are exceptions, and implement some
rudimentary prioritization based on the high level strategy. Following
available  resources and  timing,  AIUR's {\em  Spend Manager}  orders
Base,  Production   and  Construction  Managers  what   they  have  to
build/produce. It also orders to start tech research and upgrades. The
idea here is not to let the different managers allocate the resources they want, but to do the opposite, that is: finding how the AI can
spend the available money.

@@

A total of 8 bots competed in the 2013 AIIDE competition with many of the same names from the 2012 competition, and some key updates were made to existing bots for 2013 which shook up the results from the previous years. A total of 5597 games were played during the 5 day tournament. UAlbertaBot took 1st place with a dominant 84.49\% win rate, with Skynet in 2nd with 72.77\%. The major addition to UAlbertaBot was a combat simulation package called SparCraft. UAlbertaBot was re-engineered so that it now always attacked from the first combat unit created, using the SparCraft system to simulate whether or not the bot could a fight against known enemy units, retreating automatically if it thought it could not. This addition, combined with some additional bug fixes led to the victory. Skynet and AIUR both implemented strategy learning in 2013, which is evident from \ref{fig:aiide2013} in which both Skynet and AIUR's win percentage over time climb dramatically from the first few rounds of the tournament to the later rounds. Ximp unfortunately crashed all games on the Fortress map and lost 10\% of its games for free as a result, which meant it could have possibly came in 2nd place if not for all those crashes.

@@

The results of the 2014 competition were dramatically different from previous years. Since 2011 the top 3 bots in the AIIDE competition had been some permutation of Skynet, UAlbertaBot, and AIUR, and in 2014 the top 3 finishers were all completely different. In first place was IceBot with a win rate of 85.85\%, which had a complete overhaul in terms of strategy and tactics from the previous year. IceBot had several strategies which it could implement, most starting with an early bunker defense and transitioning into attacking once a set of heuristics conditions had been met. IceBot's source code was originally based on AIUR's modular design, and changed to play the Terran race. In a close 2nd place finish was Ximp - which had fixed the crashes and logic errors which had plagued it the year before. Ximp implemented a solid forge-first fast expansion strategy which transitioned into late-game carrier play. By building a large amount of early photon cannons in its expansion, Ximp was easily able to stop most of the early rushing bots, and then holding them off until its carriers cleaned up in the late game. 3rd place was taken by LetaBot, a new Terran bot whose source code was written on top of the 2012 version of UAlbertaBot and changed to play the Terran race. LetaBot implemented a number of strategies including a 'cheesy' but strong bunker rush strategy which killed many bots very quickly, as well as a Wraith (mid-game flying unit) strategy.

@@

The 2015 AIIDE competition was  the largest competition to date hosted
by a conference, with 23 total bots competing. The competition environment was also changed so that the tournament was now run on virtual machines instead of physical machines. Doing this enabled the tournament to run for a full two weeks instead of the normal one week, resulting in a total of 20788 games being played - 90 full round robins between each bot pairing. 2015 saw the most even distribution of race selection ever in a Starcraft AI competition, with 5 new Zerg submissions along with 7 Protoss bots and 9 Terran bots. The first ever Random race entry was also submitted as UAlbertaBot had been updated to play all three of the available races. When playing Random race, Starcraft randomly assigns one of Protoss, Terran, or Zerg after the game has started, with your opponent not knowing which race you are until they have scouted you on the map. This provides a significant advantage as the enemy's initial build order must now have a period of uncertainty during which is has to guess what race it is playing against.

The results of the 2015 competition were extremely close, with 1st / 2nd place as well as 3rd / 4th place being separated by less than 1\% win rate. 1st place was won by Tscmoo, a new Zerg bot which played nearly a dozen different strategies, learning which one to choose over time via the file I/O system. In a close second place was ZZZKbot, which implement a 4-pool Zergling rush strategy every game. Despite the relatively simple strategy, most bots did not have proper defense capabilities and lost in very short games. In 3rd place was Overkill, another Zerg bot which had several different strategies including Mutalisks and Hydralisks, which learned over time as well. In a close 4th place was UAlbertaBot, which played Random and implemented three main rushing strategies - one for each race. An interesting result from this tournament was that despite UAlbertaBot coming 4th place, it finished with a winning rate greater than 50\% vs each other bot. AIUR came in 5th place and was a clear demonstration of how learning over time can dramatically improve results in a tournament, going from 63\% win rate early in the competition to a final win rate of over 73\%.

\begin{table}[!t]
\caption{Results of the five best bots of the AIIDE 2015 competition}
\label{tab:aiide2015}
\centering
\begin{tabular}{|c|c|c|}
\hline
{\bfseries Rank} & {\bfseries Bot} & {\bfseries Win \%} \\
\hline
1 & Tscmoo & 88.52\% \\
2 & ZZZKBot & 87.84\% \\
3 & Overkill & 80.69\% \\
4 & UAlbertaBot & 80.20\% \\
5 & AIUR & 73.02\% \\
\hline
\end{tabular}
\end{table}

@@

The two top bots of every bracket qualified
for the final round. Table~\ref{tab:cig-first-round} summarizes
the results.
Note that as BroodwarBotQ and BTHAI have the same number of wins,
their direct encounter was evaluated which accounted 6:4 for the BroodwarBotQ.
The bots going into the final were thus UAlbertaBot, Skynet (from bracket A)
and Xelnaga and BroodwarBotQ (from bracket B). All qualified bots play the
Protoss faction. Most bots proved fairly stable, only Xelnaga and Protoss 
Beast Jelly crashed relatively often (each in more than a quarter of the games). 
Crashing of course resulted in an instant win for the other bot.
In some cases, neither bot was able to finish the other off completely,
so that they went into a passive state. Such games were manually ended after
around 15 minutes and victory were assigned to the bot that had obtained more
points as indicated on the end game screen.

@@

The final round was played in a similar mode as each of the
first round brackets,
using another set of 5 previously unknown maps:
\emph{iCCup lost temple 2.4}, \emph{iCCup rush hour 3.1},
\emph{iCCup swordinthemoon 2.1}, \emph{iCCup yellow 1.1},
and \emph{La\_Mancha 1.1}. 
Letting each pairing play on each map twice again with
switching starting positions resulted in 30 games per bot.
The final results are displayed in table~\ref{tab:cig-final-round},
indicating Skynet as winner and UAlbertaBot as runner-up, being
almost equally strong, and the two other bots as clearly inferior.
The competition setup, documentation and results can be found
in the 2011 competition web page\footnote{\url{http://ls11-www.cs.tu-dortmund.de/rts-competition/StarCraft-cig2011}}.

@@

\subsubsection{CIG 2014.}
The CIG 2014 competition was organized by Kyung-Joong Kim, Ho-Chul Cho, and In-Seok Oh of Sejong University. For 2014, the CIG competition used a total of 20 different maps which were unknown to the competitors before the competition started, which was by far the most maps ever used in a Starcraft AI competition and presented a challenge to many of the bots who entered. A total of 4680 games were played which meant that each bot played each other bot 60 times, or 3 times per map. The CIG 2014 competition was held just a few weeks before the AIIDE 2014 competition and so many of the entrants to the competition were identical, and the results definitely showed this. The top 4 bots were all the same as the 2014 AIIDE competition with IceBot coming 1st, Ximp in 2nd, LetaBot in 3rd and AIUR in 4th place. For descriptions of these bots please refer to the AIIDE 2014 competition description as none of the top-finishing bots had any significant changes made.

\begin{table}[t]
\caption{Results of the CIG 2014 competition.}
\label{tab:cig2014}
\centering
\begin{tabular}{|c|c|c|}
\hline
{\bfseries Rank} & {\bfseries Bot} & {\bfseries Win \%} \\
\hline
1 & IceBot & 83.1\% \\
2 & Ximp & 78.1\% \\
3 & LetaBot & 68.5\% \\
4 & AIUR & 66.1\% \\
5 & UAlbertaBot & 60.0\% \\ 
\hline
\end{tabular}
\end{table}

\subsubsection{CIG 2015.}
There were some significant rule changes to the CIG 2015 competition, which was once organized by members from Sejong University. The most significant rule change was that entrants no longer had to be open source, in an attempt to attract more competitors who may not want to open source their bots. The second rule change was that a single competitor could submit multiple entries to the competition - which caused some controversy among registrants since this introduces the possibility of collusion between entries. Thankfully no such collusion was detected during the competition. The 2015 competition was also not run quite as long as previous competitions with only 2730 games being played in total, or 30 between each bot pairing, 6 between each bot on each of the 5 chosen maps. There were several new Zerg entries to the CIG 2015 competition which ended up finishing in the top three positions, with results very similar to those of the AIIDE 2015 competition. ZZZBot took 1st place, tscmoo-Z (a Zerg bot written by tscmoo) came 2nd, and Overkill came 3rd. For descriptions of these bots and their strategies used please refer to the AIIDE 2015 competition section as the strategies remained largely unchanged between the CIG and AIIDE competitions.

@@

% As the figures show, the top three ranked bots in the competition (Skynet, AIUR and UalbertaBot) do not change their strategy at all depending on their opponent. For example, the Skynet bot (both in 2011 and 2012), always uses the same opening (\emph{two\_gates}), except when playing a Terran opponent, when it uses \emph{nony}. This reflects the trend that the performance of bots is still more dependent on carefully handcrafted and non-adaptive behaviors, than on on-line decision making procedures. This is so, since most of the problems that need to be solved in order to implement such procedures are still open.
